Clustering exploration for `r params$sample_id`

### 2024-09-12 analyses/cell-type-wilms-tumor-06/notebook_template/03_clustering_exploration.Rmd
change to PR#750

@@

Here, we quickly checked annotations that are present in the _processed rds object. 
However, the automated annotation have not been performed using a cancer specific reference or a kidney reference. 
We do not expect a nice labelling of the cells as the overlap of cell types between the reference and the query dataset is poor. 
This support the need to do a proper label transfer from the fetal kidney atlas, which appears to be the best reference that can be applied to a Wilms tumor query.

@@

@@

@@

In the previous step of the analysis, we selected two fetal references containing kidney samples and transfer the labels.

Here we aim to understand the overlap and differences between the two references.

##### Human fetal kidney atlas Stewart et al.

Stewart et al. created a [human fetal kidney atlas](https://www.kidneycellatlas.org/) [[3](https://www.science.org/doi/10.1126/science.aat5031)].

This reference contains only fetal kidney cells and has been precisely annotated by kidney experts.

##### Azimuth Human fetal atlas Cao et al.

`Azimuth` also provides a human fetal atlas as a reference [[4](https://www.science.org/doi/10.1126/science.aba7721)].

The data can be found on [Zenodo](https://zenodo.org/records/4738021#.YJIW4C2ZNQI).

The reference contain cells from 15 organs including kidney from fetal samples.

We were first nicely surprised that most of the cells in the Wilms tumor dataset mapped to "kidney" organ (see previous report `02a_fetal_all_reference_Cao_{sample_id}.html`).

@@

>>>>

#### Introduction to the four annotation strategies

For these reasons, we expect Wilms tumor cells to map to fetal cells, especially fetal kidney cells.

Here, we investigate annotations resulting from 4 different strategies:

* the `_processed.rds` objects already contain automated annotations computed by the [Data Lab](https://scpca.readthedocs.io/en/latest/getting_started.html):

    + Annotations determined by SingleR, an automated reference-based method [Looney et al. 2019](https://www.nature.com/articles/s41590-018-0276-y).

    + Annotations determined by CellAssign, an automated marker-gene based method [Zhang et al. 2019](https://www.nature.com/articles/s41592-019-0529-1).

* the label transfers from two fetal references using `Azimuth`, that have been performed in the previous analysis

    + Using the human fetal atlas as a reference: 
    This is the reference developed by [Cao et al.](https://www.science.org/doi/10.1126/science.aba7721) provided by `Azimuth` for label transfer.
    The reference contain cells from 15 organs including kidney from fetal samples.
    The label transfer have been performed in the notebook `02a_fetal_full_reference_Cao_{sample_id}.Rmd`

    + Using the human fetal kidney atlas: 
    [Stewart et al.](https://www.science.org/doi/10.1126/science.aat5031) created a [human fetal kidney atlas](https://www.kidneycellatlas.org/). 
    This reference contains only fetal kidney cells and has been precisely annotated by kidney experts. 
    The label transfer have been performed in the notebook `02b_fetal_kidney_reference_Stewart_{sample_id}.Rmd`

@@

@@

The fetal full reference (Cao et al.) provides three levels of annotations:

- `fetal_full_predicted.organ` is one of the 15 organs used for building the reference. 
We were nicely surprised that most of the cells from the Wilms tumor cohort maped to fetal kidney cells.

- `fetal_full_predicted.annotation.l1` holds predicted cell type annotation

- `fetal_full_predicted.annotation.l2` holds the combination of the predicted organ and predicted cell type annotation

@@

The fetal kidney reference (Stewart et al.) provides two levels of annotations:

- `fetal_kidney_predicted.compartment` is one of the 4 main compartments composing a fetal kidney.
We expect immune and endothelial cells to be healthy (non-canerous) cells identified easily with high confidency. 
We expect stroma and fetal nephron compartment to contain both normal and malignant cells.

    - `immune` cells

    - `stroma` cells

    - `fetal_nephron` cells

    - `endothelial` cells

- `fetal_full_predicted.annotation.l1` holds predicted cell type annotations that details further the four compartments, as for example:

    - `immune` subtypes: myeloid cells, lymphocytes, ...

    - `stroma` subtypes: mesenchymal stem cells, fibroblasts, ...

    - `fetal_nephron` subtypes: podocytes, uteric bud, ...

    - `endothelial` cells

@@

@@

#### Comparison of the  fetal reference with `singleR` and `cellassign`

```{r fig.width=20, fig.height=15, out.width='100%'}
p1 <- do_Table_Heatmap( srat,
                            last_group = "fetal_kidney_predicted.compartment",
                            first_group = "cellassign_celltype_annotation"

                            )
p2 <- do_Table_Heatmap( srat,
                            last_group = "fetal_kidney_predicted.compartment",
                            first_group = "singler_celltype_annotation"

```{r fig.width=20, fig.height=30, out.width='100%'}
p1 <- do_Table_Heatmap( srat,
                            last_group = "fetal_kidney_predicted.cell_type",
                            first_group = "cellassign_celltype_annotation"

p1 <- do_Table_Heatmap( srat,
                            last_group = "fetal_kidney_predicted.cell_type",
                            first_group = "singler_celltype_annotation"

p1/p2
```
